Concurrent Programming with CXL

### 2020-06-21 book/paper.tex
Latest

\begin{itemize}
\item Chapter~\ref{ch:cxlintro} introduces the CXL programming
language, as it provides the language for presenting synchronization
problems and solutions.
\item Chapter~\ref{ch:concurrent} illustrates the problem of
concurrent programming through a simple example in which two processes
are concurrently incrementing a counter.
\item Chapter~\ref{ch:cxlmachine} presents the
CXL virtual machine to understand the problem
underlying concurrency better.
\item Chapter~\ref{ch:critical} introduces the concept of a
\emph{critical section} and presents various flawed implementations
of critical sections to demonstrate that implementing a critical section
is not trivial.
\item Chapter~\ref{ch:peterson} introduces Peterson's Algorithm, an
elegant solution to implementating a critical section.
\item Chapter~\ref{ch:method} gives some more details on the CXL
language.
\item Chapter~\ref{ch:spinlock} introduces ``hardware'' locks
for implemented critical sections.
\item Chapter~\ref{ch:synch} presents CXL modules
that supports locks and other synchronization primitives.
\item Chapters~\ref{ch:rdwrbusy} and~\ref{ch:rdwr2} present the
abstraction and implementation of reader/writer
locks, which allow multiple processes in a critical section if they are
all only executing read-only accesses.
Chapter~\ref{ch:rdwrbusy} also talks about busy-waiting, which is
an undesirable approach to synchronize processes.
\item Chapter~\ref{ch:deadlock} describes \emph{deadlock}
where a set of processes are indefinitely waiting for one another to
release a resource.
\item Chapter~\ref{ch:semaphore} introduces \emph{semaphores},
a generalization of locks
that is good not only for mutual exclusion but also for waiting for
certain application-level conditions.
\item Chapters~\ref{ch:bb} and~\ref{ch:sbs} introduce the
\emph{bounded buffer problem} (aka \emph{producer/consumer problem})
and various solutions.
\item Chapter~\ref{ch:starvation} talks about \emph{starvation}:
the problem that in some
synchronization approaches processes may not be able to get access to a
resource they need.
\item Chapters~\ref{ch:monitors} and~\ref{ch:mesa} present
\emph{monitors} and \emph{condition variables},
another approach to process synchronication.
\item Chapter~\ref{ch:deadlockprevention} revisits deadlocks and how to
avoid them.
\item Chapter~\ref{ch:actor} presents the \emph{actor model} as an approach to
synchronization.
\item Chapter~\ref{ch:nonblocking} introduces \emph{non-blocking} or
\emph{wait-free} synchronization algorithms,
which prevent processes waiting for one another more than a bounded number of
steps.
\item Chapter~\ref{ch:barrier} describes \emph{barrier synchronization},
useful in high-performance computing applications such as parallel simulations.
\item Chapter~\ref{ch:abp} presents a problem and a solution to the distributed
systems problem of having two processes communicate reliably over an unreliable
network.
\end{itemize}

@@

A good example of a resource with multiple units is the
so-called \emph{bounded buffer} (aka \emph{ring buffer}).
\index{ring buffer}
It is essentially
a queue implemented using a circular buffer
\index{circular buffer}
of a certain length and two pointers:
one where items are inserted and one from where items are extracted.  If the
buffer is full, processes that want to add more items have to wait.  In the usual
formulation, processes that want to extract an item when the buffer is empty
also have to wait.
This problem is known as the ``Producer/Consumer Problem.''

@@

@@

\chapter{Mesa Condition Variables in CXL}
\label{ch:mesa}

>>>>

\begin{itemize}
\item Chapter~\ref{ch:cxlintro} introduces the CXL programming
language, as it provides the language for presenting synchronization
problems and solutions.
\item Chapter~\ref{ch:concurrent} illustrates the problem of
concurrent programming through a simple example in which two processes
are concurrently incrementing a counter.
\item Chapter~\ref{ch:cxlmachine} presents the
CXL virtual machine to understand the problem
underlying concurrency better.
\item Chapter~\ref{ch:critical} introduces the concept of a
\emph{critical section} and presents various flawed implementations
of critical sections to demonstrate that implementing a critical section
is not trivial.
\item Chapter~\ref{ch:peterson} introduces Peterson's Algorithm, an
elegant solution to implementating a critical section.
\item Chapter~\ref{ch:method} gives some more details on the CXL
language.
\item Chapter~\ref{ch:spinlock} introduces ``hardware'' locks
for implemented critical sections.
\item Chapter~\ref{ch:synch} presents CXL modules
that supports locks and other synchronization primitives.
\item Chapters~\ref{ch:rdwrbusy} and~\ref{ch:rdwr2} present the
abstraction and implementation of reader/writer
locks, which allow multiple processes in a critical section if they are
all only executing read-only accesses.
Chapter~\ref{ch:rdwrbusy} also talks about busy-waiting, which is
an undesirable approach to synchronize processes.
\item Chapter~\ref{ch:deadlock} describes \emph{deadlock}
where a set of processes are indefinitely waiting for one another to
release a resource.
\item Chapter~\ref{ch:semaphore} introduces \emph{semaphores},
a generalization of locks
that is good not only for mutual exclusion but also for waiting for
certain application-level conditions.
\item Chapters~\ref{ch:bb} and~\ref{ch:sbs} introduce the
\emph{bounded buffer problem} (aka \emph{producer/consumer problem})
and various solutions.
\item Chapter~\ref{ch:starvation} talks about \emph{starvation}:
the problem that in some
synchronization approaches processes may not be able to get access to a
resource they need.
\item Chapter~\ref{ch:monitors} presents
\emph{monitors} and \emph{condition variables},
another approach to process synchronication.
\item Chapter~\ref{ch:deadlockprevention} revisits deadlocks and how to
avoid them.
\item Chapter~\ref{ch:actor} presents the \emph{actor model} as an approach to
synchronization.
\item Chapter~\ref{ch:nonblocking} introduces \emph{non-blocking} or
\emph{wait-free} synchronization algorithms,
which prevent processes waiting for one another more than a bounded number of
steps.
\item Chapter~\ref{ch:barrier} describes \emph{barrier synchronization},
useful in high-performance computing applications such as parallel simulations.
\item Chapter~\ref{ch:abp} presents a problem and a solution to the distributed
systems problem of having two processes communicate reliably over an unreliable
network.
\end{itemize}

@@

A good example of a resource with multiple units is the
so-called \emph{bounded buffer} (aka \emph{ring buffer}).
\index{ring buffer}
Suppose you have a \emph{producer} (some process) wanting to
communicate a stream of items with a \emph{consumer} (another process).
A bounded buffer is essentially
a queue implemented using a circular buffer
\index{circular buffer}
of a certain length and two pointers:
one where the producer inserts items, and one where the consumer extracts
items.  If the
buffer is full, the producer must wait; if the buffer is empty, the
consumer must wait.
This problem is known as the ``Producer/Consumer Problem'' and was
proposed by Dijkstra~\cite{EWD329}.
Here we generalize it to allowing multiple producers and multiple consumers
all sharing the same bounded buffer.

@@

\section*{Exercises}
\begin{enumerate}
\item TBD.  starvation-free dressing room
\item TBD.  multiple GPU allocator
\end{enumerate}

@@

% \chapter{Mesa Condition Variables in CXL}
% \label{ch:mesa}
